---
title: "Introduction"
author: "Céline Trébeau"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css


## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=TRUE}
library("DNABarcodeCompatibility")
write.table(DNABarcodeCompatibility::IlluminaIndexes, file <- tempfile(), row.names = FALSE, col.names = FALSE, quote=FALSE)
barcodes <- file_loading_and_checking(file)
head(barcodes)
head(get_all_combinations(barcodes, 3, 4))

m <- get_random_combinations(barcodes, 3, 4)
filtered_m <- distance_filter(barcodes, m, "hamming", 3)
optimize_combinations(filtered_m, 12, 48)

experiment_design(file1=file, sample_number=18, mplex_level=3, chemistry=4)

```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
