---
title: "Simulations"
author: "Trébeau, Boutet de Monvel, et al., Etournay."
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  BiocStyle::pdf_document:
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=120)
# options(digits=3)

SCRIPT_DIR=file.path(Sys.getenv("HOME"),"DNABarcodeCompatibility/vignettes/")
if (!file.exists(SCRIPT_DIR)) stop("Script directory not properly set up !")
library("dplyr")
library("ggplot2")
```

```{r style, echo = FALSE, results = 'asis', warnings=FALSE, messages=FALSE, eval=F}
BiocStyle::markdown()
```

# Introduction

DNABarcodeCompatibility works in 2 steps:

* Step1: find compatible combinations of barcodes

* Step2: optimize the set of compatible barcodes to reduce redundancy between barcodes
    + Step2a: optimization by the greedy search method
    + Step2b: optimization by N iterations of the greedy search method


# Simulations relative to Step1 of the workflow: find compatible combinations of barcodes

To find compatible combinations of DNA barcodes, one approach is to generate all possible combinations of DNA barcodes and perform an exhaustive search of compatible combinations.
However, the total number of combinations may become prohibitively large for large barcode sets and large multiplexing numbers. Thus, an exhaustive search of compatible combinations among all possible combinations may not be possible. An alternive strategy is to choose a fixed number of randomly selected combinations for finding compatible combinations. To determine the conditions for which a random search is preferable over an exhaustive search, we will consider the execution time of the program as a critical parameter. 


## Set parameters for the simulations

The execution time of the program depends on the size of the barcode dataset, the multiplex level and the Illumina chemistry. We therefore test different values of these parameters to vizualise how they influence the execution time.

```{r echo = FALSE, cache=F, results=F}

R=10
L = c(12,18,24,30,36,42,48)
N = 2:5
C = c(1,2,4)

# R=2
# L = c(12,18)
# N = 2:5
# C = c(2,4)

print("CONDITIONS:")
print("Barcode set size:")
L
print("Multiplex levels:")
N 
print("Chemistry:")
C
print("Number of simulations per condition:")
R
print(paste("Total number of simulations:",R*length(L)*length(N)*length(C)))

```

Number of simulations per condition | Barcode-set size | Multiplex level | Chemistry |
------------------------------------|------------------|-----------------|-----------|
`r R`                               |`r L`             | `r N`           | `r C`     |


Number of simulations | |
----------------------|-|
`r R*length(L)*length(N)*length(C)`| |


## Run simulations for exhaustive searches
```{r echo = FALSE, cache=F}

# Initialize output file1 (barcode occurrences)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), mplex_level=integer(0), barcode_number=integer(0), chemistry=integer(0)),
 outputfile1 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file2 (execution time)
write.table(data.frame(rep_number=integer(0), time=numeric(0), nb_total_combinations=integer(0), nb_comp_combinations=integer(0), mplex_level=integer(0), barcode_number=integer(0), chemistry=integer(0)),
 outputfile2 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Define Illumina TruSeq small RNA barcode set
inputfile <- tempfile()
save(DNABarcodeCompatibility::IlluminaIndexes, file=inputfile)

# Detect CPU number
nb_cpu=parallel::detectCores()

if (F) {
  # Run simulation using GNU parallel (Windows users need cygwin + Gnu parallel; MacOs: brew install parallel; Linux: apt-get install parallel)
  print(paste("Running", R*length(L)*length(N)*length(C) ,"simulations..."))
  system.time({
    system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_get_all_combinations.R {1} {2} {3} {4} {5} {6} {7} ::: ",paste(1:R, collapse=" ")," ::: ",paste(N, collapse=" ")," ::: ",paste(L, collapse=" ")," ::: ",paste(C, collapse=" "), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(inputfile, collapse="")))
  })
  
  barcodeFreq <- read.table(file=outputfile1, header = T, sep = " ", stringsAsFactors = FALSE)
  save(barcodeFreq, file = paste0(SCRIPT_DIR,"barcodeFreq"))

  results <- read.table(file=outputfile2, header = T, sep = " ", stringsAsFactors = FALSE)
  save(results, file = paste0(SCRIPT_DIR,"results"))
}


# Import results
barcodeFreq <- local(get(load(paste0(SCRIPT_DIR,"barcodeFreq"))))
results <- local(get(load(paste0(SCRIPT_DIR,"results"))))

```



### Execution time of an exhaustive search as a function of the total number of combinations

The execution time of an exhaustive search increases as a power of the total number of combinations. It is slighty faster for the 4-channel chemistry. 
For the ease of use, we fix to a few seconds the time of execution. On a standard computer, this corresponds to an exhaustive search among about 2000 combinations for the 4-channel chemistry.
Above 2000 combinations, the program will then perform a search of compatible barcodes from a randomly generated set of barcode combinations. Other important aspects also motivate this choice (see below).


```{r echo = FALSE, fig.width=11, fig.wide=T}
results %>% group_by(mplex_level, barcode_number, chemistry) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations)) %>%
  ungroup() %>% mutate(mplex_level=as.factor(mplex_level),
                       channel=paste0(chemistry, "-channel")) %>%
ggplot(aes( x = nb_total_combinations, y = time.avg))+
  geom_point()+
  geom_errorbar(aes(ymin=(time.avg-time.sd), ymax=(time.avg+time.sd)),
                size=0.7,color="black") +
  scale_x_log10(breaks= c(10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7), labels=scales::scientific) +
  scale_y_log10(breaks = c(1, 2, 5, 10,30,90,200, 1200, 3600, 5000))+
  geom_hline(yintercept=30, linetype="dashed", color = "red")+
  geom_hline(yintercept=10, linetype="dashed", color = "red")+
  geom_hline(yintercept=2, linetype="dashed", color = "red")+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~channel) +
  ggtitle(paste("Execution time for increasing number of combinations ( average over", R,"simulations )"))+
  xlab("Total number of combinations")+
  ylab("Execution time [s]")
```


### Execution time of an exhaustive search as a function of the barcode-set size

The total number of combinations can be expressed as a binomial coefficient ${n \choose k}$ that is indexed by $n$ barcodes in the barcode dataset and a mutliplex level $k$, where both $n$ and $k$ are positive integers.
We now plot the execution time as function of the barcode-set size for various multiplex levels. This reveals that the execution time is less than 2 seconds for a multiplex level of 2 with the 48-barcode Illumina small RNA TruSep kits.
Therefore, in these conditions, it's very advantageous to perform an exhaustive search of compatible combinations for a multiplex level of 2. For a multiplex level higher than 4 and large barcode sets, we however expect the program to perform a search of compatible barcodes from a randomly generated set of barcode combinations.


```{r echo = FALSE, fig.width=11}
results %>% group_by(mplex_level, barcode_number, chemistry) %>%
  # filter(chemistry > 1) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations)) %>%
  ungroup() %>% mutate(mplex_level=as.factor(mplex_level),
                       channel=paste0(chemistry, "-channel")) %>%
ggplot(aes( x = barcode_number, y = time.avg, color = mplex_level))+
  geom_point()+
  geom_errorbar(aes(ymin=(time.avg-time.sd), ymax=(time.avg+time.sd)),
                size=0.3,color="black") +
  scale_x_log10(breaks=L) +
  scale_y_log10(breaks = c(1, 2,5, 10,30,90,200,1000,5000))+
  geom_hline(yintercept=30, linetype="dashed", color = "red")+
  geom_hline(yintercept=10, linetype="dashed", color = "red")+
  geom_hline(yintercept=2, linetype="dashed", color = "red")+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~channel) +
  ggtitle(paste("Execution time by multiplex levels ( average over", R,"simulations )"))+
  xlab("Barcode set size")+
  ylab("Execution time [s]")


```


### Probability to find compatible barcodes in the complete ensemble of combinations

* The probability to find compatible combinations is more variable for small randomly-selected barcode sets 

```{r echo = FALSE, warning=FALSE, fig.width=11}
results %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>% filter(mplex_level==3) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations), nb_comp_combinations.sd=sd(nb_comp_combinations)) %>%
  ungroup() %>% mutate(mplex_level=as.factor(mplex_level),
                       channel=paste0(chemistry, "-channel"),
                       comp_comb_prop.avg= nb_comp_combinations.avg/nb_total_combinations,
                       comp_comb_prop.sd= nb_comp_combinations.sd/nb_total_combinations,
                       barcode_number=as.factor(barcode_number)) %>%
  ggplot(aes(x=barcode_number,y=comp_comb_prop.avg, fill = mplex_level))+
  geom_histogram(stat = "identity")+
  geom_errorbar(aes(ymin=(comp_comb_prop.avg-comp_comb_prop.sd), ymax=(comp_comb_prop.avg+comp_comb_prop.sd)),
                size=0.3,color="black") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))+
  facet_grid(~channel) +
  scale_fill_manual(values = c("darkgrey")) + 
  # scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1)) +
  xlab("Barcode set size")+
  ylab("Proportion of compatible combinations")+
ggtitle(paste("Average proportion of compatible combinations ( average over", R,"simulations )"))


```


* The probability to find compatible combinations increases with the multiplex level


```{r echo = FALSE, warning=FALSE, fig.width=11}
results %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>%
  summarise(nb_total_combinations=nb_total_combinations[1], nb_comp_combinations=mean(nb_comp_combinations), comp_comb_prop=nb_comp_combinations/nb_total_combinations) %>%
  group_by(mplex_level, chemistry) %>% 
  summarize(comp_comb_prop.avg= mean(comp_comb_prop),
            comp_comb_prop.sd= sd(comp_comb_prop)) %>% ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         channel=paste0(chemistry, "-channel")) %>%
  ggplot(aes(x=mplex_level,y=comp_comb_prop.avg, fill = mplex_level))+
  geom_histogram(stat = "identity")+
  geom_errorbar(aes(ymin=(comp_comb_prop.avg-comp_comb_prop.sd), ymax=(comp_comb_prop.avg+comp_comb_prop.sd)),
  size=0.3,color="black") +
  # theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))+
  facet_grid(~channel) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1)) +
  xlab("Multiplex level")+
  ylab("Proportion of compatible combinations")+
  ggtitle(paste("Probability to find compatible combinations ( average over", R,"simulations and barcode-set size )"))


```

* Relationship between the number of compatibles combinations and the total number of combinations 

For a multiplex level of 2, the number of compatible combinations is less than 100 for the combined Illumina TruSeq small RNA kits. Therefore, for a multiplex level of 2, an exhaustive search is preferable to retrieve all possible compatible combinations of barcodes for the optimization steps. We fix to 2024 the total number of combinations beyond which it's however quicker and sufficient to perform a random search of compatible combinations (see dashed line in the graph below). This number is motivated by two reasons. Below this threshold, an exhaustive search will be systematically performed for a multiplex level of 2 (less than 2024 combinations in total). Above this threshold, a random search will provide enough compatible combinations for futher optimization steps. Indeed, the above results indicate that the probability to get a compatible combination is at least 18% for multiplex levels >= 3. Briefly, the random search will randomly pick up an arbitrarily fixed number of 1000 combinations representing a pool of at least 180 compatible combinations, which is suitable for further optimization steps (see below). Thus, the `DNABarcodeCompatibility::experiment_design()` function of the package will use a threshold of 2024 combinations in total to switch between an exhaustive search and a random search of compatible combinations.



```{r echo = FALSE, warning=FALSE, fig.width=11}
results %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>%
  summarise(nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations),
            nb_comp_combinations.sd=sd(nb_comp_combinations)) %>%
  ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         channel=paste0(chemistry, "-channel")) %>%
  ggplot(aes(x=nb_total_combinations,y=nb_comp_combinations.avg, color = mplex_level, size=barcode_number))+
  geom_point()+ geom_smooth(method= "lm", se=F) +
  geom_vline(xintercept=2024, linetype="dashed", color = "red")+
  scale_x_log10(breaks=c(10,100,2024,10^4,10^5, 10^6), labels = scales::comma) +
  scale_y_log10(breaks=c(5,10,100,1000,10^4,10^5, 10^6), labels = scales::comma) +
  scale_size_continuous(name = "Barcode-set size", breaks=unique(results$barcode_number)) +
  facet_grid(~channel) +
  xlab("Total number of combinations [log scale]")+
  ylab("Number of compatible combinations [log scale]")+
  ggtitle(paste("Number of compatible combinations in the Illumina TruSeq small RNA kits"))


```

### Compatible barcode combinations: average barcode frequencies among input barcode sets

The distribution of barcodes used to form compatible combinations may be highly heterogeneous at low multiplex levels. 
We plot the distribution of barcodes averaged over the different barcode sets, for distinct chemistries and multiplex levels.
Despite the averaging of the distibutions over different barcode sets of different sizes, the heterogeneity between barcode frequencies is clearly visible.

* 4-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (4 %in% barcodeFreq$chemistry) {
  barcodeFreq %>% filter(chemistry == 4) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (4-channel chemistry)"))
}
```

* 2-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (2 %in% barcodeFreq$chemistry) {
  barcodeFreq %>% filter(chemistry == 2) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (2-channel chemistry)"))
}
```

* 1-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (1 %in% barcodeFreq$chemistry) {
  barcodeFreq %>% filter(chemistry == 1) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (1-channel chemistry)"))
}

```

### Compatible barcode combinations: single examples

Individual barcode sets of 12 barcodes were randomly generated from the four Illumina TruSeq small RNA kits comprising 48 barcodes in total. We clearly see that barcode occurences are very heterogeneous even for multiplex levels of 4 or 5. Therefore, consumable usage may strongly be biased among preparation kits.

* 4-channel chemistry, multiplex level 4, barcode-set size 12

```{r echo = FALSE, warning=FALSE, fig.width=8}
chem=4
mpl=4
bcnb=12
repnb=3

nb_comp_combinations <- (results %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$nb_comp_combinations

barcodeFreq %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurrences")+
ggtitle(paste("Barcode occurences among",nb_comp_combinations,"compatible barcode combinations (",chem,"- channel )"))
```

* 4-channel chemistry, multiplex level 5, barcode-set size 12

```{r echo = FALSE, warning=FALSE, fig.width=8}
chem=4
mpl=5
bcnb=12
repnb=3

nb_comp_combinations <- (results %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$nb_comp_combinations

barcodeFreq %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurences")+
ggtitle(paste("Barcode occurences among",nb_comp_combinations,"compatible barcode combinations (",chem,"- channel )"))
```


* 2-channel chemistry, multiplex level 4, barcode-set size 12

```{r echo = FALSE, warning=FALSE, fig.width=8}
chem=2
mpl=4
bcnb=12
repnb=3

nb_comp_combinations <- (results %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$nb_comp_combinations

barcodeFreq %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurences")+
ggtitle(paste("Barcode occurences among",nb_comp_combinations,"compatible barcode combinations (",chem,"- channel )"))
```



# Simulations relative to Step2 of the workflow

* Step2: optimize the set of compatible barcodes to minimize redundancy between barcodes

The goal is to estimate the probability to get an optimal solution and to test how the optimizer improves the solution compared with a random pick. 

Here, we compare two greedy search algorithms that we implemented for this project, the greedy descent and the greedy exchange algorithms. We use different subsets of randomly-generated compatible combinations, and for each of those we perform 100 simulations with both algorithms, in order to estimate the probability $P_{1}$ to get an optimal solution. This optimal solution consists of a set of compatible barcode combinations in which the barcode distribution is uniform or close to uniform. The number of barcode combinations in the optimal solution is defined by the product of the multiplex level by the number of lanes, which defines the number of libraries to be sequenced.

Based on this probability $P_{1}$, we devise an optimizer function that performs $N$ trials of the greedy search. The probability to get an optimal solution after $N$ iterations reads $P_{N} = 1-(1-P_{1})^N$. For example, if we want to find an optimal solution at least 95% of cases $(P_{N} = 0.95)$, we need a number of iterations $N \geq log(1-P_{N})/log(1-P_{1})$.  


* We therefore run two series of simulations:

**Step2a:** find the probability $P_{1}$ of to get an optimal combination given the subset size, and estimate $N$ for $P_{N} = 0.95$, in order to refine the optimization in a reasonable computation time.


**Step2b:** check the efficacy of the optimizer function on a few challenging examples


## Step2a: set parameters for the simulations


```{r echo = FALSE, results=F}
R=100
L = c(12,18,24,30)
N = 2:5
C = c(2,4)
thrs_size_comb_for_GD=c(seq(10,50,5),seq(60,100, 10))
thrs_size_comb_for_GE=c(seq(10,50,5),seq(60,100, 10), seq(150,300,50), 400)
max_iteration = 0
nb_lane=c(4,6,8)
algorithm= c("greedy_exchange", "greedy_descent")

# R=20
# L = c(18)
# N = 4
# C = c(4)
# thrs_size_comb_for_GD=c(60)
# thrs_size_comb_for_GE=c(160)
# max_iteration = 0
# nb_lane=c(4)
# algorithm= c("greedy_exchange", "greedy_descent")


print("CONDITIONS:")
print("Barcode set size:")
N
print("Multiplex levels:")
L 
print("Chemistry:")
C
print("Number of simulations per greedy-descent search:")
R
print("Size of compatible combinations set used for greedy-descent:")
thrs_size_comb_for_GD

print("Size of compatible combinations set used for greedy-exchanges:")
thrs_size_comb_for_GE

print("Number of iterations for the greedy-descent:")
max_iteration

print("Number of lanes:")
nb_lane

print("Total number of combinations:")
choose(L,N)

print("Optimizer algorithms")
algorithm

print(paste("Number of conditions:",R*length(L)*length(N)*length(C)*(length(thrs_size_comb_for_GD)+length(thrs_size_comb_for_GE))*length(nb_lane)))

```

Number of simulations per condition | Barcode-set size | Multiplex level | Chemistry | Lane number | Algorithm | Compatible combinations (GD) | Compatible combinations (GE) |
------------------------------------|------------------|-----------------|-----------|----------------|--------------|---------------------------|---------------------------|
`r R`                               |`r L`             | `r N`           | `r C`     | `r nb_lane`    | `r algorithm`| `r thrs_size_comb_for_GD` | `r thrs_size_comb_for_GE` |


Number of simulations | |
----------------------|-|
`r R*length(L)*length(N)*length(C)*(length(thrs_size_comb_for_GD)+length(thrs_size_comb_for_GE))*length(nb_lane)`| |


## Run simulations for step2a


```{r echo = FALSE, cache=F}

# Initialize output file1 (barcode occurrence from random pic)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile1 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file2 (barcode occurrence from optimized set)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile2 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file3 (calcul of proba to get an optimal solution)
write.table(data.frame(time=numeric(0), nb_comp_combinations=integer(0), rep_number=integer(0), barcode_number=integer(0), thrs_size_comb=integer(0), S_max=numeric(0), S_init=numeric(0), S_random=numeric(0), S_opt=numeric(0), S_random_opt=numeric(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile3 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file4 (barcode occurrence from optimized set based on random picks)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile4 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Detect CPU number
nb_cpu=parallel::detectCores()

if (F) {
  system.time({
    for (i in L) {
      # Generate a large set of compatible combinations
      index <- dplyr::sample_n(DNABarcodeCompatibility::IlluminaIndexes, nrow(DNABarcodeCompatibility::IlluminaIndexes), replace = FALSE)
      for (j in N) {
        for (k in C) {
          for (algo in algorithm) {
            
            # Initialize input file with a matrix of all combinations
            combination_m <- DNABarcodeCompatibility::get_all_combinations(index_df = index[1:i,], mplex_level = j, chemistry = k)
            
            inc=0
            while (nrow(combination_m) < max(nb_lane) && inc<10) {
              print("nrow(combination_m) might be too small")
              print("other trial...")
              index <- dplyr::sample_n(DNABarcodeCompatibility::IlluminaIndexes, nrow(DNABarcodeCompatibility::IlluminaIndexes), replace = FALSE)
              combination_m <- DNABarcodeCompatibility::get_all_combinations(index_df = index[1:i,], mplex_level = j, chemistry = k)
              inc=inc+1
              print(inc)
            }
            if (nrow(combination_m)<max(nb_lane)) {warning("nrow(combination_m) might be too small")}
            
            inputfile <- tempfile()
            save(combination_m, file=inputfile)
            # save(combination_m, file="combination_m_4ch")
            
            # Run simulation using GNU parallel (Windows users need cygwin + Gnu parallel; MacOs: brew install parallel; Linux: apt-get install parallel)
            if (algo == "greedy_descent") {
              system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_optimize_combinations.R {1} {2} {3} {4} {5} {6} {7} {8} {9} {10} {11} {12} {13} ::: ",paste(1:R, collapse=" ")," ::: ",paste(thrs_size_comb_for_GD, collapse=" ")," ::: " , paste(max_iteration, collapse=" "), " ::: ",paste(inputfile, collapse=""), " ::: ", paste(nb_lane, collapse=" "), " ::: ", paste(i, collapse=""), " ::: ", paste(k, collapse=""), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(outputfile3, collapse=""), " ::: ", paste(outputfile4, collapse=""), " ::: ", paste(j, collapse=""), " ::: ", paste(algo, collapse=" ")))
            } else if (algo == "greedy_exchange") {
              system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_optimize_combinations.R {1} {2} {3} {4} {5} {6} {7} {8} {9} {10} {11} {12} {13} ::: ",paste(1:R, collapse=" ")," ::: ",paste(thrs_size_comb_for_GE, collapse=" ")," ::: " , paste(max_iteration, collapse=" "), " ::: ",paste(inputfile, collapse=""), " ::: ", paste(nb_lane, collapse=" "), " ::: ", paste(i, collapse=""), " ::: ", paste(k, collapse=""), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(outputfile3, collapse=""), " ::: ", paste(outputfile4, collapse=""), " ::: ", paste(j, collapse=""), " ::: ", paste(algo, collapse=" ")))
            }
          }
        }
      }
    }
    
    sim1_barcodeStat_random <- read.table(file=outputfile1, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim1_barcodeStat_random, file = paste0(SCRIPT_DIR,"sim1_barcodeStat_random"))
    
    sim1_barcodeStat_opt <- read.table(file=outputfile2, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim1_barcodeStat_opt, file = paste0(SCRIPT_DIR,"sim1_barcodeStat_opt"))
    
    sim1_probaSmax <- read.table(file=outputfile3, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim1_probaSmax, file = paste0(SCRIPT_DIR,"sim1_probaSmax"))
    
    sim1_barcodeStat_random_opt <- read.table(file=outputfile4, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim1_barcodeStat_random_opt, file = paste0(SCRIPT_DIR,"sim1_barcodeStat_random_opt"))
  })
}


# Import results
sim1_barcodeStat_random <- local(get(load(paste0(SCRIPT_DIR,"sim1_barcodeStat_random"))))
sim1_barcodeStat_opt <- local(get(load(paste0(SCRIPT_DIR,"sim1_barcodeStat_opt"))))
# sim1_barcodeStat_random_opt <- local(get(load(paste0(SCRIPT_DIR,"sim1_barcodeStat_random_opt"))))
sim1_probaSmax <- local(get(load(paste0(SCRIPT_DIR,"sim1_probaSmax"))))
```



```{r echo = FALSE, fig.width=11, eval=F}
for (i in thrs_size_comb){
  # for (j in barcode_number) {
    stat <- sim1_probaSmax %>% filter(thrs_size_comb==i) %>%
      select(-c(rep_number,barcode_number,thrs_size_comb,chemistry,nb_lane, mplex_level)) %>%
      summary()
    print(i)
    print(stat)
  # }
}
```
### Execution time of a single greedy search run for different sizes of compatible combination subsets

We perform `r R` trials of the greedy search for several conditions (see table above). When the set of compatible combinations is large, the computation time may become prohibitively large. To reduce the computation time, we randomly generate a subset of compatible combinations that we use to perform the optimization by a greedy search approach. Here, we want to know how the size of the subset influences the computation time for each of the greedy search algorithm.

We averaged the execution time over `r R` replicates of different combinations of conditions of chemistry, barcode-set size, number of lanes, and multiplex level. Part of the variability observed for the execution time comes from the diversity of conditions that may substantially influence the execution time.

The graph below indicates that the running time of the greedy-descent algorithm is quadratic as a function of the barcode-set size unlike the greedy-exchange, which is linear.

```{r echo = FALSE,  warning=FALSE, message=F}
# In linear coordinates
sim1_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>%
  group_by(algorithm, thrs_size_comb) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1]) %>%
  ungroup() %>% 
ggplot(aes( x = thrs_size_comb, y = time.avg, color=algorithm)) +
  geom_point() + geom_line() + geom_smooth() +
  geom_errorbar(aes(ymin=(time.avg-time.sd), ymax=(time.avg+time.sd)),
                color="black") +
   scale_color_manual("Algorithm",
                       breaks = c("greedy_descent", "greedy_exchange"),
                       values = c("blue", "red")) +
  # geom_smooth(method = "lm", se = FALSE)+
  # scale_x_continuous(breaks=c(10, 50, 100, 200, 300, 400)) +
  # scale_y_continuous(breaks = c(1, 15,30,90,200, 250))+
  # geom_hline(yintercept=30, linetype="dashed", color = "red")+
  # geom_hline(yintercept=15, linetype="dashed", color = "red")+
  # geom_hline(yintercept=1, linetype="dashed", color = "red")+
  # facet_wrap(~algorithm) +
  ggtitle(paste("Execution time of a single greedy-descent: averaged for", R, "runs")) +
  xlab("Size of the initial subset of compatible combinations") +
  ylab("Execution time [s]")

```

```{r echo = FALSE,  warning=FALSE, message=F, fig.width=11, fig.height=22, eval=F}
# In log-log coordinates
sim1_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1]) %>%
  ungroup() %>% 
  mutate(mplex_level=as.factor(mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
          barcode_number=paste("set of",barcode_number,"barcodes")) %>%
ggplot(aes( x = thrs_size_comb, y = time.avg, color=mplex_level)) +
  geom_point() + #geom_line() + geom_smooth() +
  geom_errorbar(aes(ymin=(time.avg-time.sd), ymax=(time.avg+time.sd)),
                color="black") +
  geom_smooth(method = "lm", se = T)+
  scale_x_log10(breaks=c(10, 50, 80, 100, 200, 300)) +
  scale_y_log10(breaks = c(1, 5,15,30,90,200, 250))+
  geom_hline(yintercept=30, linetype="dashed", color = "red")+
  geom_hline(yintercept=15, linetype="dashed", color = "red")+
  geom_hline(yintercept=1, linetype="dashed", color = "red")+
  facet_grid(chemistry+nb_lane+algorithm~barcode_number) +
  ggtitle(paste("Execution time of a single greedy-descent: averaged for", R, "runs")) +
  xlab("Size of the initial subset of compatible combinations") +
  ylab("Execution time [s]")

```


### Probability to get an optimal set of barcode combinations (max entropy) for a single greedy search

We perform `r R` trials of the greedy search for several conditions (see table above) to estimate the probability to reach the theoretical maximum entropy $S_{max}$ in each condition. Here, we test how the size of the subset influences the probability to reach $S_{max}$.

To our surprise, results are very heterogeneous among conditions. However one common trend is that larger subsets of compatible combinations (> 50 compatible combinations) give better results.
In some cases, $S_{max}$ is not even reached. Either the probability $P_{1}$ is below 1% or there is not any set of compatible combination that can found. 

```{r echo = FALSE,  warning=FALSE, message=F, fig.width=11, fig.height=22}

sim1_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>%
  mutate(isSmax=ifelse(abs(S_opt-S_max)<10^(-10), T, F)) %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1], S_max=S_max[1], ProbaOfSmax=sum(isSmax)/length(rep_number)) %>%
  ungroup() %>% 
  mutate(mplex_level=as.factor(mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) %>%
ggplot(aes( x = thrs_size_comb, y = ProbaOfSmax, color=mplex_level)) +
  geom_point() +
  geom_smooth(se = F) +
  scale_x_continuous(breaks=seq(0, 400, 50)) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1.1)) +
  facet_grid(chemistry+nb_lane+algorithm~barcode_number) +
  
  ggtitle(paste("Proba to get exactly Smax after a single run: calculated from", R, "runs")) +
  xlab("Size of the initial set of compatible combinations") +
  ylab("Probability to get an optimal combination")

```

### Probability to get the best optimized set of barcode combinations for a single greedy search

In some cases, the probability to reach the theoretical maximum entropy $S_{max}$ might be too low or null. Instead, we estimate the probability to get the highest value of the entropy $S_{opt}$ where possibly $S_{opt} < S_{max}$. In these conditions, the barcode distribution won't necessarily be uniform, but still more optimized than by a random pick (see below).


```{r echo = FALSE,  warning=FALSE, message=F, fig.width=11, fig.height=22}

sim1_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  mutate(S_sup=max(S_opt),
         isSsup=ifelse(abs(S_opt-S_sup)<10^(-10), T, F)) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1], S_sup=S_sup[1], ProbaOfSsup=sum(isSsup)/length(rep_number)) %>%
  ungroup() %>% 
  mutate(mplex_level=as.factor(mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) %>%
ggplot(aes( x = thrs_size_comb, y = ProbaOfSsup, color=mplex_level)) +
  geom_point() +
  geom_smooth(se = F) +
  scale_x_continuous(breaks=seq(0, 400, 50)) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1.1)) +
  facet_grid(chemistry+nb_lane+algorithm~barcode_number) +
  
  ggtitle(paste("Proba to get the best optimized set after a single run: calculated from", R, "runs")) +
  xlab("Size of the initial set of compatible combinations") +
  ylab("Probability to get an optimal combination")

```



### Estimation of the maximum execution time (N iterations) of a greedy search to obtain the max entropy in 95% of cases


 The probability of getting an optimal solution after $N$ iterations reads $P_{N} = 1-(1-P_{1})^N$. For example, if we want to find an optimal solution at least 95% of cases $(P_{N} = 0.95)$, we need a number of iterations $N \geq log(1-P_{N})/log(1-P_{1})$.
 

* Max execution time / iteration number as a function of the number of compatible barcodes used for the search, grouped by chemistry, algorithm, barcode-set size, and multiplex level

The results shown in the graphs below are heterogeneous between conditions. However common trends emerge. First, in general, the greedy-exchange search performs the optimization faster over N iterations than the greedy-descent search, except for a few cases (see next point) where the subset of compatible combinations is smaller than 30. Second, for compatible-combination subsets larger than 120 (only tested for the greedy-exchange), the optimization requires less iterations in general and performs faster. Based on these observations, it seems reasonable to fix the compatible-combination subset size to 120.

```{r echo = FALSE,  warning=FALSE, fig.width=11, fig.height=22}
# for 95% of cases
P=0.95

sim1_probaSmax %>% 
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>% 
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  mutate(S_sup=max(S_opt),
         isSsup=ifelse(abs(S_opt-S_sup)<10^(-10), T, F)) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1], S_sup=S_sup[1], ProbaOfSsup=sum(isSsup)/length(rep_number)) %>%
  ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) %>%
  mutate(N=ceiling(log(1-P)/log(1-ProbaOfSsup)),
         exec_time=(N+1)*time.avg) %>%
  ggplot(aes(x = thrs_size_comb, y = exec_time, color=N, shape=mplex_level)) +
  geom_point(size = 2) +
  scale_colour_gradientn(colours = rainbow(5)) +
  scale_x_log10(breaks=c(10, 15, 20, 40,60, 80,120, 200, 300)) +
  scale_y_log10(breaks = c(1, 2,5, 10,30,90,200,1000,5000,10000, 20000)) +
  geom_hline(yintercept=30, linetype="dashed", color = "black")+
  geom_hline(yintercept=10, linetype="dashed", color = "black")+
  geom_hline(yintercept=2, linetype="dashed", color = "black")+
  facet_grid(chemistry+nb_lane+algorithm~barcode_number) +
  ggtitle(paste("Estimated execution time to find an optimal solution in",P*100,"% of cases")) +
  xlab("Size of the initial subset of compatible combinations") +
  ylab("Maximum execution time [s]")


```

* Max execution time /iteration number as a function of lane number, grouped by algorithm, multiplex level, and number of compatible barcodes used for the search

Data are now pooled regardless chemistry and barcode-set sizes. From the 2 graphs below, we confirm that the greedy-descent search performs faster for small subsets of compatible combinations.
For subsets larger than 30 compatible combinations, it seems that 200 iterations would be a reasonable number for an efficient and fast optimization by the greedy-exchange search.

```{r echo = FALSE,  warning=FALSE, fig.width=11, fig.height=11}
# for 95% of cases
P=0.95

sim1_probaSmax %>% 
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>% 
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  mutate(S_sup=max(S_opt),
         isSsup=ifelse(abs(S_opt-S_sup)<10^(-10), T, F)) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_comp_combinations=nb_comp_combinations[1], S_sup=S_sup[1], ProbaOfSsup=sum(isSsup)/length(rep_number)) %>%
  mutate(isNbLibSupToNbbc=ifelse(mplex_level*nb_lane>barcode_number, T, F),
         isNbCombInfToNbLane=ifelse(nb_comp_combinations<nb_lane, T, F)) %>%
  ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) %>%
  mutate(N=ceiling(log(1-P)/log(1-ProbaOfSsup)),
         exec_time=(N+1)*time.avg) %>%
  filter(thrs_size_comb %in% c(10, 30, 80, 100, 400)) -> dat


dat %>%
  ggplot(aes(x = as.factor(nb_lane), y = N, color=mplex_level)) + 
  geom_boxplot() +
  facet_grid(thrs_size_comb~algorithm) +
  ggtitle(paste("Distribution of the number of iterations needed to find a best optimized solution in",P*100,"% of cases")) +
  xlab("Number of lanes") +
  ylab("Number of greedy-search iterations")


dat %>%
  ggplot(aes(x = as.factor(nb_lane), y = exec_time, color=mplex_level)) +
  geom_boxplot() +
  scale_y_log10(breaks=c(2, 10, 30, 100, 360, 3600)) +
  geom_hline(yintercept=30, linetype="dashed", color = "darkgrey")+
  geom_hline(yintercept=10, linetype="dashed", color = "darkgrey")+
  geom_hline(yintercept=2, linetype="dashed", color = "darkgrey")+
  facet_grid(thrs_size_comb~algorithm) +
  ggtitle(paste("Estimated execution time to find a best optimized solution in",P*100,"% of cases")) +
  xlab("Number of lanes") +
  ylab("Maximum execution time [s]")


```

### Distributions of entropy values without (S_random) and with (S_opt) optimization

To visualize the efficacy of the greedy search algorithms, we compare the distribution of entropy values obtained from randomly generated sets of compatible barcodes with the distribution of entropy values obtained from optimized sets of compatible barcodes. The graphs below show that both greedy algorithms are comparable and dramatically reduce the barcode redanduncy by favoring sets of compatible barcode combinations with a higher entropy. One multifacetted graph is shown per subset size. The blue line indicates the position of the theoretical value of the maximum entropy $S_{max}$. A dashed line indicates that $S_{max}$ has been reached, whereas a solid line indicates that $S_{max}$ has not been reached.

```{r echo = FALSE,  warning=FALSE, message=FALSE, fig.width=11, fig.height=44}
sim1_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>% ungroup() %>%
  # filter(thrs_size_comb == 60) %>%
  mutate(isSmax=ifelse(abs(S_opt-S_max)<10^-10, T, F),
         S_opt=round(S_opt, 3),
         S_random=round(S_random, 3), 
         mplex_level=paste0("mplex-",mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) -> dat

dat %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  do ({
    chunk <- .
    with(chunk, as.data.frame(table(S=S_random))) %>% 
      mutate(Freq=Freq/sum(Freq),
             S_max=chunk$S_max[1],
             dataset="S_random",
             isSmax=any(chunk$isSmax))
  }) %>% ungroup() %>% mutate(S=as.double(S)) -> S_random_freq 


dat %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  do ({
    chunk <- .
    with(chunk, as.data.frame(table(S=S_opt))) %>% 
      mutate( Freq=Freq/sum(Freq), 
              S_max=chunk$S_max[1], 
              dataset="S_opt",
              isSmax=any(chunk$isSmax)) 
  }) %>% ungroup() %>% mutate(S=as.double(S)) -> S_opt_freq


Sdat=bind_rows(S_random_freq, S_opt_freq)

k=1
gg <- list()

for (i in c(10,30,80,100,400) ){
  
  Sdat %>% filter(thrs_size_comb == i) %>%
    # mutate(thrs_size_comb=as.factor(thrs_size_comb)) %>%
    ggplot(aes(x = S)) +
    geom_histogram(aes(y=0.05*..density.., fill=dataset), position='identity', alpha=0.7, binwidth=0.05, boundary=0.05) +
    geom_vline(aes(xintercept = S_max, color="S_max", linetype=isSmax)) +
    scale_x_continuous(breaks=round(seq(min(dat$S_random), max(dat$S_opt), (max(dat$S_opt)-min(dat$S_random))/5),2)) +
    # scale_x_discrete(breaks=sort(sample(dat$S_random,5)))+
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    facet_grid(chemistry+mplex_level+nb_lane+algorithm~barcode_number) +
    scale_fill_manual("Entropy",
                      breaks = c("S_opt", "S_random"),
                      values = c("green", "red")) +
    
    scale_color_manual("",
                       breaks = c("S_max", "S_opt", "S_random"),
                       values = c("blue", "green", "red")) +
    xlab("Entropy values") +
    ylab("Frequency") +
    ggtitle(paste("Distributions of entropy values for (",i ,"combs;", R, "runs)")) -> gg[[k]]
  k=k+1
}
gg

```


```{r echo = FALSE,  warning=FALSE, message=FALSE, fig.width=11, eval=F}
if (640 %in% sim1_probaSmax$thrs_size_comb) {
  sim1_probaSmax %>% mutate(isSmax=ifelse(S_opt==S_max, T, F)) %>%
    filter(thrs_size_comb == 640) -> dat
  
  dat %>%
    mutate(thrs_size_comb=as.factor(thrs_size_comb)) %>%
    ggplot(aes(y = (..count..)/sum(..count..))) +
   geom_histogram(aes(x = S_opt,fill="S_opt"), alpha=0.5, center=dat$S_max[1],  binwidth = 0.01) +
    geom_histogram(aes(x = S_random,fill="S_random"), alpha=0.5, center=dat$S_max[1], binwidth = 0.01) +
    # geom_density(aes(x = S_opt,fill="S_opt"), alpha=0.5, color= "black", adjust=0.1) +
    # geom_density(aes(x = S_random,fill="S_random"), alpha=0.5, color= "black", adjust=0.1) +
    geom_vline(aes(xintercept = S_max[1], color="S_max"), linetype="dashed") +
    scale_fill_manual("Entropy",
                      breaks = c("S_opt", "S_random"),
                      values = c("green", "red")) +
    scale_color_manual("",
                       breaks = c("S_max", "S_opt", "S_random"),
                       values = c("blue", "green", "red")) +
    facet_wrap(~thrs_size_comb) +
    xlab("Entropy values") +
    ylab("Frequency") +
    ggtitle(paste("Frequency distributions of entropy values for", R, "runs")) 
}
```


## Step2b: set parameters for the simulations

The results of the simulations in step2a suggest a few cases for which the optimization procedure would require more than 100 iterations to reach $S_{max}$. We find interesting to use these challenging conditions to test how good our greedy-exchange algorithm would perform compared to a basic optimization by iterative random picks. In the table below are summarized a few challenging conditions to be tested.

```{r echo = FALSE, results=F}
R=100
L = c(12, 30)
N = 4:5
C = c(4)
thrs_size_comb_for_GE=c(120)
thrs_size_comb_for_GD=""
max_iteration = 200
nb_lane=c(6)
algorithm="greedy_exchange"

# R=1
# L = c(18)
# N = 4
# C = c(4)
# thrs_size_comb=c(80)
# max_iteration = 14
# nb_lane=c(4)


print("CONDITIONS:")
print("Barcode set size:")
N
print("Multiplex levels:")
L 
print("Chemistry:")
C
print("Number of simulations per greedy-descent search:")
R
print("Size of compatible combinations set used for greedy-descent:")
thrs_size_comb_for_GE

print("Number of iterations for the greedy-descent:")
max_iteration

print("Number of lanes:")
nb_lane

print("Total number of combinations:")
choose(L,N)

print(paste("Number of conditions:",R*length(L)*length(N)*length(C)*length(thrs_size_comb_for_GE)*length(nb_lane)))

```


Number of simulations per condition | Barcode-set size | Multiplex level | Chemistry | Lane number | Algorithm | Compatible combinations (GD) | Compatible combinations (GE) | Max iterations N |
------------------------------------|------------------|-----------------|-----------|----------------|--------------|---------------------------|---------------------------|----------------|
`r R`                               |`r L`             | `r N`           | `r C`     | `r nb_lane`    | `r algorithm`| `r thrs_size_comb_for_GD` | `r thrs_size_comb_for_GE` | `r max_iteration` |


Number of simulations | |
----------------------|-|
`r R*length(L)*length(N)*length(C)*length(thrs_size_comb_for_GE)*length(nb_lane)`| |


## Run simulations for step2b

```{r echo = FALSE, cache=T}

# Reuse the dataset generated in the first series of simulations
combination_m_4ch <- local(get(load(paste0(SCRIPT_DIR,"combination_m_4ch"))))

inputfile <- tempfile()
save(combination_m_4ch, file=inputfile)

# Initialize output file1 (barcode occurrence from random pic)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile1 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file2 (barcode occurrence from optimized set)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile2 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file3 (calcul of proba to get an optimal solution)
write.table(data.frame(time=numeric(0), nb_comp_combinations=integer(0), rep_number=integer(0), barcode_number=integer(0), thrs_size_comb=integer(0), S_max=numeric(0), S_init=numeric(0), S_random=numeric(0), S_opt=numeric(0), S_random_opt=numeric(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile3 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file4 (barcode occurrence from optimized set based on random picks)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), thrs_size_comb=integer(0), barcode_number=integer(0), chemistry=integer(0), nb_lane=integer(0), mplex_level=integer(0), algorithm=character(0)),
            outputfile4 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Detect CPU number
nb_cpu=parallel::detectCores()

if (F) {
  # Run simulation using GNU parallel (Windows users need cygwin + Gnu parallel; MacOs: brew install parallel; Linux: apt-get install parallel)
  system.time({
 system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_optimize_combinations.R {1} {2} {3} {4} {5} {6} {7} {8} {9} {10} {11} {12} {13} ::: ",paste(1:R, collapse=" ")," ::: ",paste(thrs_size_comb_for_GE, collapse=" ")," ::: " , paste(max_iteration, collapse=" "), " ::: ",paste(inputfile, collapse=""), " ::: ", paste(nb_lane, collapse=" "), " ::: ", paste(i, collapse=""), " ::: ", paste(k, collapse=""), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(outputfile3, collapse=""), " ::: ", paste(outputfile4, collapse=""), " ::: ", paste(j, collapse=""), " ::: ", paste(algorithm, collapse=" ")))
  })
  
  sime2_barcodeStat_random <- read.table(file=outputfile1, header = T, sep = " ", stringsAsFactors = FALSE)
  save(sime2_barcodeStat_random, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_random"))
  
  sim2_barcodeStat_opt <- read.table(file=outputfile2, header = T, sep = " ", stringsAsFactors = FALSE)
  save(sim2_barcodeStat_opt, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_opt"))
  
  sim2_probaSmax <- read.table(file=outputfile3, header = T, sep = " ", stringsAsFactors = FALSE)
  save(sim2_probaSmax, file = paste0(SCRIPT_DIR,"sim2_probaSmax"))
  
  sim2_barcodeStat_random_opt <- read.table(file=outputfile4, header = T, sep = " ", stringsAsFactors = FALSE)
  save(sim2_barcodeStat_random_opt, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_random_opt"))
}


# Detect CPU number
nb_cpu=parallel::detectCores()

if (F) {
  system.time({
    for (i in L) {
      # Generate a large set of compatible combinations
      index <- dplyr::sample_n(DNABarcodeCompatibility::IlluminaIndexes, nrow(DNABarcodeCompatibility::IlluminaIndexes), replace = FALSE)
      for (j in N) {
        for (k in C) {
          for (algo in algorithm) {
            
            # Initialize input file with a matrix of all combinations
            combination_m <- DNABarcodeCompatibility::get_all_combinations(index_df = index[1:i,], mplex_level = j, chemistry = k)
            
            inc=0
            while (nrow(combination_m) < max(nb_lane) && inc<10) {
              print("nrow(combination_m) might be too small")
              print("other trial...")
              index <- dplyr::sample_n(DNABarcodeCompatibility::IlluminaIndexes, nrow(DNABarcodeCompatibility::IlluminaIndexes), replace = FALSE)
              combination_m <- DNABarcodeCompatibility::get_all_combinations(index_df = index[1:i,], mplex_level = j, chemistry = k)
              inc=inc+1
              print(inc)
            }
            if (nrow(combination_m)<max(nb_lane)) {warning("nrow(combination_m) might be too small")}
            
            inputfile <- tempfile()
            save(combination_m, file=inputfile)
            # save(combination_m, file="combination_m_4ch")
            
            # Run simulation using GNU parallel (Windows users need cygwin + Gnu parallel; MacOs: brew install parallel; Linux: apt-get install parallel)
            if (algo == "greedy_descent") {
              system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_optimize_combinations.R {1} {2} {3} {4} {5} {6} {7} {8} {9} {10} {11} {12} {13} ::: ",paste(1:R, collapse=" ")," ::: ",paste(thrs_size_comb_for_GD, collapse=" ")," ::: " , paste(max_iteration, collapse=" "), " ::: ",paste(inputfile, collapse=""), " ::: ", paste(nb_lane, collapse=" "), " ::: ", paste(i, collapse=""), " ::: ", paste(k, collapse=""), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(outputfile3, collapse=""), " ::: ", paste(outputfile4, collapse=""), " ::: ", paste(j, collapse=""), " ::: ", paste(algo, collapse=" ")))
            } else if (algo == "greedy_exchange") {
              system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_optimize_combinations.R {1} {2} {3} {4} {5} {6} {7} {8} {9} {10} {11} {12} {13} ::: ",paste(1:R, collapse=" ")," ::: ",paste(thrs_size_comb_for_GE, collapse=" ")," ::: " , paste(max_iteration, collapse=" "), " ::: ",paste(inputfile, collapse=""), " ::: ", paste(nb_lane, collapse=" "), " ::: ", paste(i, collapse=""), " ::: ", paste(k, collapse=""), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(outputfile3, collapse=""), " ::: ", paste(outputfile4, collapse=""), " ::: ", paste(j, collapse=""), " ::: ", paste(algo, collapse=" ")))
            }
          }
        }
      }
    }
    
    sim2_barcodeStat_random <- read.table(file=outputfile1, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim2_barcodeStat_random, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_random"))
    
    sim2_barcodeStat_opt <- read.table(file=outputfile2, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim2_barcodeStat_opt, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_opt"))
    
    sim2_probaSmax <- read.table(file=outputfile3, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim2_probaSmax, file = paste0(SCRIPT_DIR,"sim2_probaSmax"))
    
    sim2_barcodeStat_random_opt <- read.table(file=outputfile4, header = T, sep = " ", stringsAsFactors = FALSE)
    save(sim2_barcodeStat_random_opt, file = paste0(SCRIPT_DIR,"sim2_barcodeStat_random_opt"))
  })
}  

        
# Import results
sim2_barcodeStat_random <- local(get(load(paste0(SCRIPT_DIR,"sim2_barcodeStat_random"))))
sim2_barcodeStat_random_opt <- local(get(load(paste0(SCRIPT_DIR,"sim2_barcodeStat_random_opt"))))
sim2_barcodeStat_opt <- local(get(load(paste0(SCRIPT_DIR,"sim2_barcodeStat_opt"))))
sim2_probaSmax <- local(get(load(paste0(SCRIPT_DIR,"sim2_probaSmax"))))

```

###  Distributions of entropy values after running the optimizer

We compare the distributions of entropy values obtained using 3 different strategies over at max `r max_iteration` iterations per trial. We performed `r R` trials for each strategy. We calculate and compare the 3 resulting entropy values:

1. **S_random**: no optimization; a subset of compatible combinations is randomly picked up only once (no iteration)
2. **S_opt**: a subset of compatible combinations is optimized using the greedy-exchange algorithm (`r max_iteration` iterations at max)
3. **S_random_opt**: a subset of compatible combinations is optimized by random pick using the same amount of iterations used with the greedy search.

* Set of `r thrs_size_comb_for_GE` compatible combinations and up to `r max_iteration` iterations per trial

In the graphs below, the blue line indicates the position of the theoretical value of the maximum entropy $S_{max}$. Solid lines indicate that $S_{max}$ has not been reached over `r max_iteration` iterations. Either the probability to find $S_{max}$ is too low to be reached in `r max_iteration` iterations, or  $S_{max}$ cannot be reached due to constraints in the barcode set. Furthermore, the dashed line indicates that $S_{max}$ has been reached.

In any case, the greedy search algorithm performs better than a random search, and converges to the best entropy value.

```{r echo = FALSE,  warning=FALSE, message=FALSE, fig.width=11, fig.height=11}
sim2_probaSmax %>% group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm) %>%
  filter(thrs_size_comb<nb_comp_combinations) %>% ungroup() %>%
  # filter(thrs_size_comb == 60) %>%
  mutate(isSmax=ifelse(abs(S_opt-S_max)<10^-10, T, F),
         S_opt=round(S_opt, 3),
         S_random=round(S_random, 3), 
         mplex_level=paste0("mplex-",mplex_level),
         chemistry=paste0(chemistry,"-channel"),
         nb_lane=paste(nb_lane,"lanes"),
         barcode_number=paste("set of",barcode_number,"barcodes")) -> dat

dat %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  do ({
    chunk <- .
    with(chunk, as.data.frame(table(S=S_random))) %>% 
      mutate(Freq=Freq/sum(Freq),
             S_max=chunk$S_max[1],
             dataset="S_random",
             isSmax=any(chunk$isSmax))
  }) %>% ungroup() %>% mutate(S=as.double(S)) -> S_random_freq 


dat %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  do ({
    chunk <- .
    with(chunk, as.data.frame(table(S=S_opt))) %>% 
      mutate( Freq=Freq/sum(Freq), 
              S_max=chunk$S_max[1], 
              dataset="S_opt",
              isSmax=any(chunk$isSmax)) 
  }) %>% ungroup() %>% mutate(S=as.double(S)) -> S_opt_freq

dat %>%
  group_by(chemistry, barcode_number, mplex_level, nb_lane, algorithm, thrs_size_comb) %>%
  do ({
    chunk <- .
    with(chunk, as.data.frame(table(S=S_random_opt))) %>% 
      mutate( Freq=Freq/sum(Freq), 
              S_max=chunk$S_max[1], 
              dataset="S_random_opt",
              isSmax=any(chunk$isSmax)) 
  }) %>% ungroup() %>% mutate(S=as.double(S)) -> S_random_opt_freq


Sdat=bind_rows(S_random_freq, S_opt_freq, S_random_opt_freq)


gg <- list()
k=1
for (i in unique(thrs_size_comb_for_GE,thrs_size_comb_for_GD) ){
  
  Sdat %>% filter(thrs_size_comb == i) %>%
    # filter(isSmax) %>%
    # mutate(thrs_size_comb=as.factor(thrs_size_comb)) %>%
    ggplot(aes(x = S)) +
    geom_histogram(aes(y=0.05*..density.., fill=dataset), position='identity', alpha=0.6, binwidth=0.050, boundary=0.05) +
    geom_vline(aes(xintercept = S_max, color="S_max", linetype=isSmax)) + #, linetype="dashed"
    scale_x_continuous(breaks=round(seq(min(dat$S_random), max(dat$S_opt), (max(dat$S_opt)-min(dat$S_random))/5),2)) +
    # scale_x_discrete(breaks=sort(sample(dat$S_random,5)))+
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    facet_grid(chemistry+mplex_level+nb_lane+algorithm~barcode_number) +
    scale_fill_manual("Entropy",
                      breaks = c("S_opt", "S_random", "S_random_opt"),
                      values = c("green", "red", "orange")) +
    
    scale_color_manual("",
                       breaks = c("S_max", "S_opt", "S_random", "S_random_opt"),
                       values = c("blue", "green", "red", "orange")) +
    xlab("Entropy values") +
    ylab("Frequency") +
    ggtitle(paste("Distributions of entropy values for (",i ,"combs;", R, "runs, up to",max_iteration,"interations )")) -> gg[[k]]
  k=k+1
}
gg

```

* Average execution time per condition

| Conditions | Set of 12 barcodes | Set of 30 barcodes |
|------------|--------------------|--------------------|
| mplex-4    | `r round(mean((sim2_probaSmax %>% filter(barcode_number==12) %>% filter(mplex_level==4))$time))` sec    |  `r round(mean((sim2_probaSmax %>% filter(barcode_number==30) %>% filter(mplex_level==4))$time))` sec |
| mplex-5    |`r round(mean((sim2_probaSmax %>% filter(barcode_number==12) %>% filter(mplex_level==5))$time))` sec    |  `r round(mean((sim2_probaSmax %>% filter(barcode_number==30) %>% filter(mplex_level==5))$time))` sec |

### Comparison of barcode distributions between a randomly generated set and an optimized set of compatible combinations

#### Case A: 30 barcodes, $S_{max}$ is reached

* Barcode set:
```{r echo = FALSE, warning=FALSE}
barcodeset <- sort(unique((sim2_barcodeStat_random_opt %>% filter(barcode_number==30) %>% filter(mplex_level==4))$barcodeID))
```

| Barcode set | |
|-------------|-|
`r barcodeset`| |


* Set of compatible combinations that was randomly selected

```{r echo = FALSE, warning=FALSE, fig.width=11}
chem=4
mpl=4
bcnb=30

# Find one repetition number for which S_random is minimum
repnb <- (sim2_probaSmax %>%
            filter(chemistry==chem) %>%
            filter(mplex_level==mpl) %>%
            filter(barcode_number==bcnb) %>%
            filter(S_random==min(S_random)))$rep_number[1]




S_random <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_random

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_random %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurrences")+
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_random=",round(S_random,2),"; S_max=",round(S_max,2),")"))

```


* Set of compatible barcodes that was optimized with the iterative greedy-exchange search

We use the same experimental conditions as above for a fair comparison.

```{r echo = FALSE, warning=FALSE, fig.width=11}
S_opt <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_opt

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_opt %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID") +
  ylab("Barcode occurrences") +
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_opt=",round(S_opt,2),"; S_max=",round(S_max,2),")"))
```


#### Case B: 30 barcodes, $S_{max}$ is **not** reached

* Barcode set:
```{r echo = FALSE, warning=FALSE}
barcodeset <- sort(unique((sim2_barcodeStat_random_opt %>% filter(barcode_number==30) %>% filter(mplex_level==5))$barcodeID))
```

| Barcode set | |
|-------------|-|
`r barcodeset`| |


* Set of compatible combinations that was randomly selected

```{r echo = FALSE, warning=FALSE, fig.width=11}
chem=4
mpl=5
bcnb=30

# Find one repetition number for which S_random is minimum
repnb <- (sim2_probaSmax %>%
            filter(chemistry==chem) %>%
            filter(mplex_level==mpl) %>%
            filter(barcode_number==bcnb) %>%
            filter(S_random==min(S_random)))$rep_number[1]




S_random <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_random

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_random %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurrences")+
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_random=",round(S_random,2),"; S_max=",round(S_max,2),")"))

```


* Set of compatible barcodes that was optimized with the iterative greedy-exchange search

We use the same experimental conditions as above for a fair comparison.

```{r echo = FALSE, warning=FALSE, fig.width=11}
S_opt <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_opt

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_opt %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID") +
  ylab("Barcode occurrences") +
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_opt=",round(S_opt,2),"; S_max=",round(S_max,2),")"))
```

#### Case C: 12 barcodes, $S_{max}$ is reached

* Barcode set:
```{r echo = FALSE, warning=FALSE}
barcodeset <- sort(unique((sim2_barcodeStat_random_opt %>% filter(barcode_number==12) %>% filter(mplex_level==5))$barcodeID))
```

| Barcode set | |
|-------------|-|
`r barcodeset`| |


* Set of compatible combinations that was randomly selected

```{r echo = FALSE, warning=FALSE, fig.width=11}
chem=4
mpl=5
bcnb=12

# Find one repetition number for which S_random is minimum
repnb <- (sim2_probaSmax %>%
            filter(chemistry==chem) %>%
            filter(mplex_level==mpl) %>%
            filter(barcode_number==bcnb) %>%
            filter(S_random==min(S_random)))$rep_number[1]




S_random <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_random

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_random %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurrences")+
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_random=",round(S_random,2),"; S_max=",round(S_max,2),")"))

```


* Set of compatible barcodes that was optimized with the iterative greedy-exchange search

We use the same experimental conditions as above for a fair comparison.

```{r echo = FALSE, warning=FALSE, fig.width=11}
S_opt <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_opt

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_opt %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID") +
  ylab("Barcode occurrences") +
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_opt=",round(S_opt,2),"; S_max=",round(S_max,2),")"))
```


#### Case D: 12 barcodes, $S_{max}$ is **not** reached

* Barcode set:
```{r echo = FALSE, warning=FALSE}
barcodeset <- sort(unique((sim2_barcodeStat_random_opt %>% filter(barcode_number==12) %>% filter(mplex_level==4))$barcodeID))
```

| Barcode set | |
|-------------|-|
`r barcodeset`| |


* Set of compatible combinations that was randomly selected

```{r echo = FALSE, warning=FALSE, fig.width=11}
chem=4
mpl=4
bcnb=12

# Find one repetition number for which S_random is minimum
repnb <- (sim2_probaSmax %>%
            filter(chemistry==chem) %>%
            filter(mplex_level==mpl) %>%
            filter(barcode_number==bcnb) %>%
            filter(S_random==min(S_random)))$rep_number[1]




S_random <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_random

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_random %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  # scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID")+
  ylab("Barcode occurrences")+
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_random=",round(S_random,2),"; S_max=",round(S_max,2),")"))

```


* Set of compatible barcodes that was optimized with the iterative greedy-exchange search

We use the same experimental conditions as above for a fair comparison.

```{r echo = FALSE, warning=FALSE, fig.width=11}
S_opt <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_opt

S_max <- (sim2_probaSmax %>% filter(chemistry == chem) %>% 
   filter(mplex_level==mpl) %>%
   filter(barcode_number==bcnb) %>%
   filter(rep_number==repnb))$S_max

sim2_barcodeStat_opt %>% filter(chemistry == chem) %>% 
  filter(mplex_level==mpl) %>%
  filter(barcode_number==bcnb) %>%
  filter(rep_number==repnb) %>%
  mutate(channel=paste0(chemistry, "-channel"),
         mplex_level=paste("mutiplex level",mplex_level)) %>%
  ggplot(aes(x=barcodeID,y=occurrence, fill=mplex_level))+
  geom_histogram(stat = "identity", fill= "darkgrey")+
  scale_y_continuous(breaks = 1:4, limits = c(0,4))+
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
  facet_grid(mplex_level~channel) +
  xlab("BarcodeID") +
  ylab("Barcode occurrences") +
ggtitle(paste0("Barcode occurences from a random set of compatible combinations (S_opt=",round(S_opt,2),"; S_max=",round(S_max,2),")"))
```

# Simulations on a randomly generated set of barcodes

## Generating a barcode set with the `DNABarcodes` package from Tilo Buschmann

We use the `DNABarcodes` package from Tilo Buschmann to generate a new set of 6-mer barcodes, which are robust against at least one substitution error (Hamming distance of 3). In addition, barcodes including stretches of 3 or more nucleotides (homopolymer of length >=3) are discarded as well as those that have an unbalanced ratio of bases G or C versus A or T. We then exclude barcodes that are present both in the Illumina TruSeq small RNA kits and in the newly generated set.

```{r echo = TRUE, warning=FALSE}
library("DNABarcodes")
newBarcodesRaw <- create.dnabarcodes(n=6, dist=3, metric = "hamming", filter.triplets = F, filter.gc = T, filter.self_complementary = F)

newBarcodes <- newBarcodesRaw[!(newBarcodesRaw %in% DNABarcodeCompatibility::IlluminaIndexes$sequence)]

ranBarcodeSet <- data.frame(Id=paste0("B",sprintf("%02d",1:length(newBarcodes))), sequence=newBarcodes, GC_content=50, homopolymer=FALSE, stringsAsFactors = F)[1:48,]
```

## Set parameters for the simulation

```{r echo = FALSE, cache=F, results=F}

R=10
L = c(12,18,24,30,36,42,48)
N = 2:5
C = c(1,2,4)

# R=2
# L = c(12,18)
# N = 2:5
# C = c(2,4)

print("CONDITIONS:")
print("Barcode set size:")
L
print("Multiplex levels:")
N 
print("Chemistry:")
C
print("Number of simulations per condition:")
R
print(paste("Total number of simulations:",R*length(L)*length(N)*length(C)))

```

Number of simulations per condition | Barcode-set size | Multiplex level | Chemistry |
------------------------------------|------------------|-----------------|-----------|
`r R`                               |`r L`             | `r N`           | `r C`     |


Number of simulations | |
----------------------|-|
`r R*length(L)*length(N)*length(C)`| |


## Run simulations for exhaustive searches
```{r echo = FALSE, cache=F}

# Initialize output file1 (barcode occurrences)
write.table(data.frame(barcodeID=character(0), occurrence=integer(0), rep_number=integer(0), mplex_level=integer(0), barcode_number=integer(0), chemistry=integer(0)),
 outputfile1 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Initialize output file2 (execution time)
write.table(data.frame(rep_number=integer(0), time=numeric(0), nb_total_combinations=integer(0), nb_comp_combinations=integer(0), mplex_level=integer(0), barcode_number=integer(0), chemistry=integer(0)),
 outputfile2 <- tempfile(), row.names = FALSE, col.names = T, quote=FALSE)

# Define Illumina TruSeq small RNA barcode set
inputfile <- tempfile()
save(ranBarcodeSet, file=inputfile)


# Detect CPU number
nb_cpu=parallel::detectCores()

if (T) {
  # Run simulation using GNU parallel (Windows users need cygwin + Gnu parallel; MacOs: brew install parallel; Linux: apt-get install parallel)
  print(paste("Running", R*length(L)*length(N)*length(C) ,"simulations..."))
  system.time({
    system(paste0("parallel -j",nb_cpu," Rscript ",SCRIPT_DIR,"sim_get_all_combinations.R {1} {2} {3} {4} {5} {6} {7} ::: ",paste(1:R, collapse=" ")," ::: ",paste(N, collapse=" ")," ::: ",paste(L, collapse=" ")," ::: ",paste(C, collapse=" "), " ::: ", paste(outputfile1, collapse=""), " ::: ", paste(outputfile2, collapse=""), " ::: ", paste(inputfile, collapse="")))
  })
  
  barcodeFreq2 <- read.table(file=outputfile1, header = T, sep = " ", stringsAsFactors = FALSE)
  save(barcodeFreq2, file = paste0(SCRIPT_DIR,"barcodeFreq2"))

  results2 <- read.table(file=outputfile2, header = T, sep = " ", stringsAsFactors = FALSE)
  save(results2, file = paste0(SCRIPT_DIR,"results2"))
}


# Import results
barcodeFreq2 <- local(get(load(paste0(SCRIPT_DIR,"barcodeFreq2"))))
results2 <- local(get(load(paste0(SCRIPT_DIR,"results2"))))

```

### Probability to find compatible barcodes in the complete ensemble of combinations

* The probability to find compatible combinations is more variable for small randomly-selected barcode sets 

```{r echo = FALSE, warning=FALSE, fig.width=11}
results2 %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>% filter(mplex_level==3) %>%
  summarise(time.avg=mean(time), time.sd=sd(time), nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations), nb_comp_combinations.sd=sd(nb_comp_combinations)) %>%
  ungroup() %>% mutate(mplex_level=as.factor(mplex_level),
                       channel=paste0(chemistry, "-channel"),
                       comp_comb_prop.avg= nb_comp_combinations.avg/nb_total_combinations,
                       comp_comb_prop.sd= nb_comp_combinations.sd/nb_total_combinations,
                       barcode_number=as.factor(barcode_number)) %>%
  ggplot(aes(x=barcode_number,y=comp_comb_prop.avg, fill = mplex_level))+
  geom_histogram(stat = "identity")+
  geom_errorbar(aes(ymin=(comp_comb_prop.avg-comp_comb_prop.sd), ymax=(comp_comb_prop.avg+comp_comb_prop.sd)),
                size=0.3,color="black") +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))+
  facet_grid(~channel) +
  scale_fill_manual(values = c("darkgrey")) + 
  # scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1)) +
  xlab("Barcode set size")+
  ylab("Proportion of compatible combinations")+
ggtitle(paste("Average proportion of compatible combinations ( average over", R,"simulations )"))


```


* The probability to find compatible combinations increases with the multiplex level


```{r echo = FALSE, warning=FALSE, fig.width=11}
results2 %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>%
  summarise(nb_total_combinations=nb_total_combinations[1], nb_comp_combinations=mean(nb_comp_combinations), comp_comb_prop=nb_comp_combinations/nb_total_combinations) %>%
  group_by(mplex_level, chemistry) %>% 
  summarize(comp_comb_prop.avg= mean(comp_comb_prop),
            comp_comb_prop.sd= sd(comp_comb_prop)) %>% ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         channel=paste0(chemistry, "-channel")) %>%
  ggplot(aes(x=mplex_level,y=comp_comb_prop.avg, fill = mplex_level))+
  geom_histogram(stat = "identity")+
  geom_errorbar(aes(ymin=(comp_comb_prop.avg-comp_comb_prop.sd), ymax=(comp_comb_prop.avg+comp_comb_prop.sd)),
  size=0.3,color="black") +
  # theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))+
  facet_grid(~channel) +
  scale_y_continuous(breaks=seq(0, 1, 0.2), limit=c(0, 1)) +
  xlab("Multiplex level")+
  ylab("Proportion of compatible combinations")+
  ggtitle(paste("Probability to find compatible combinations ( average over", R,"simulations and barcode-set size )"))


```

* Relationship between the number of compatibles combinations and the total number of combinations 

For a multiplex level of 2, the number of compatible combinations is less than 100 for the combined Illumina TruSeq small RNA kits. Therefore, for a multiplex level of 2, an exhaustive search is preferable to retrieve all possible compatible combinations of barcodes for the optimization steps. We fix to 2024 the total number of combinations beyond which it's however quicker and sufficient to perform a random search of compatible combinations (see dashed line in the graph below). This number is motivated by two reasons. Below this threshold, an exhaustive search will be systematically performed for a multiplex level of 2 (less than 2024 combinations in total). Above this threshold, a random search will provide enough compatible combinations for futher optimization steps. Indeed, the above results indicate that the probability to get a compatible combination is at least 18% for multiplex levels >= 3. Briefly, the random search will randomly pick up an arbitrarily fixed number of 1000 combinations representing a pool of at least 180 compatible combinations, which is suitable for further optimization steps (see below). Thus, the `DNABarcodeCompatibility::experiment_design()` function of the package will use a threshold of 2024 combinations in total to switch between an exhaustive search and a random search of compatible combinations.



```{r echo = FALSE, warning=FALSE, fig.width=11}
results2 %>% group_by(mplex_level, barcode_number, chemistry) %>%
  filter(chemistry > 0) %>%
  summarise(nb_total_combinations=nb_total_combinations[1], nb_comp_combinations.avg=mean(nb_comp_combinations),
            nb_comp_combinations.sd=sd(nb_comp_combinations)) %>%
  ungroup() %>%
  mutate(mplex_level=as.factor(mplex_level),
         channel=paste0(chemistry, "-channel")) %>%
  ggplot(aes(x=nb_total_combinations,y=nb_comp_combinations.avg, color = mplex_level, size=barcode_number))+
  geom_point()+ geom_smooth(method= "lm", se=F) +
  geom_vline(xintercept=2024, linetype="dashed", color = "red")+
  scale_x_log10(breaks=c(10,100,2024,10^4,10^5, 10^6), labels = scales::comma) +
  scale_y_log10(breaks=c(5,10,100,1000,10^4,10^5, 10^6), labels = scales::comma) +
  scale_size_continuous(name = "Barcode-set size", breaks=unique(results2$barcode_number)) +
  facet_grid(~channel) +
  xlab("Total number of combinations [log scale]")+
  ylab("Number of compatible combinations [log scale]")+
  ggtitle(paste("Number of compatible combinations in the Illumina TruSeq small RNA kits"))


```

### Compatible barcode combinations: average barcode frequencies among input barcode sets

The distribution of barcodes used to form compatible combinations may be highly heterogeneous at low multiplex levels. 
We plot the distribution of barcodes averaged over the different barcode sets, for distinct chemistries and multiplex levels.
Despite the averaging of the distibutions over different barcode sets of different sizes, the heterogeneity between barcode frequencies is clearly visible.

* 4-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (4 %in% barcodeFreq2$chemistry) {
  barcodeFreq2 %>% filter(chemistry == 4) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (4-channel chemistry)"))
}
```

* 2-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (2 %in% barcodeFreq2$chemistry) {
  barcodeFreq2 %>% filter(chemistry == 2) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (2-channel chemistry)"))
}
```

* 1-channel chemistry

```{r echo = FALSE, warning=FALSE, fig.width=11}
if (1 %in% barcodeFreq2$chemistry) {
  barcodeFreq2 %>% filter(chemistry == 1) %>% 
    group_by(mplex_level, chemistry, barcode_number) %>%
    mutate(sum_barcodes=sum(occurrence)) %>%
    group_by(barcodeID, mplex_level, chemistry, barcode_number) %>%
    summarise(freq=sum(occurrence)/sum_barcodes[1]) %>% 
    group_by(barcodeID, mplex_level, chemistry) %>%
    summarize(freq.avg=mean(freq), freq.sd=sd(freq)) %>% ungroup() %>%
    mutate(channel=paste0(chemistry, "-channel"),
           mplex_level=as.factor(mplex_level)) %>%
    ggplot(aes(x=barcodeID,y=freq.avg, fill=mplex_level))+
    geom_histogram(stat = "identity")+
    geom_errorbar(aes(ymin=(freq.avg-freq.sd), ymax=(freq.avg+freq.sd)),
                  size=0.7,color="black") +
    theme(axis.text.x=element_text(angle=90, hjust=1, vjust=1))+
    scale_y_continuous(breaks=seq(0, 0.18, 0.04), limit=c(0, 0.18)) +
    facet_grid(mplex_level~channel) +
    xlab("BarcodeID")+
    ylab("Barcodes frequency")+
    ggtitle(paste("Barcode frequency averaged over barcode sets (1-channel chemistry)"))
}

```

